---
title: "AntWeb"
author: "KE LeVan"
date: "January 13, 2016"
output: html_document
---

# Exploring patterns of ant occurrence data in AntWeb
Ants are a diverse group with representation in most parts of the globe. I'm interested in using the AntWeb.org api to explore ant populations. 

### Setting up libraries, paths and useful functions
```{r Libraries,warning=FALSE,echo=FALSE}
myPath <- "REPLACE with YOUR path to the repo"
myPathToData <- paste(myPath,"data",sep="/")

library(XML)
library(AntWeb)

multipleCombine <- function(input, ply = llply){
  require(plyr,warn.conflicts = FALSE)
  require(dplyr,warn.conflicts = FALSE)
  ply(input, function(x){
    t <- read.table(x, header=TRUE, sep=",",stringsAsFactors = FALSE) # read the csv
    t1 <- rbind(t) # rbind it to a temporary variable
    return(t1) # return the full variable
  }
  )
}
```

## Getting all the names of ant species
As a shortcut, I'm using AntBase as my definitive list of species. At some later date, it might be nice to check their list against ITIS.gov ('Integrated Taxonomic Information Service') - but this is good enough to get started.
```{r Compiling a complete list of named ant species}
# Scraping AntBase for a complete list global of Ant species
antBase <- "http://osuc.biosci.ohio-state.edu/hymenoptera/tsa.list_valid_taxa?query=Formicidae+Species"
allAnts <- unique.data.frame(xmlToDataFrame(getNodeSet(xmlParse(antBase,encoding = "UTF-8"),"//Species"),
                                            stringsAsFactors = FALSE))
colnames(allAnts) <- "scientificName"
```
According to AntBase, there are `r length(allAnts$scientificName)` valid ant species globally. 

## Querying the AntWeb Database for distributions of ants
```{r Getting distributions for those species}
# Querying AntWeb by scientific name
dBase <- data.frame()
  for (i in 1:dim(allAnts)[1]){
    tmp <- try(dBase <- aw_data(scientific_name = allAnts$scientificName[i])$data)
    if(is.null(tmp)==FALSE){
      dBase$scientific_name <- paste(dBase$genus,dBase$specificEpithet,sep=" ")
      write.csv(dBase,paste(myPathToData,paste(allAnts$scientificName[i],"csv",sep="."),sep="/"),
                row.names=FALSE)
    }
    if(class(tmp)=="try-error"){
      doOverTaxa <- c(doOverTaxa,i)
    }
    dBase <- data.frame()
  }

# If you don't finish in one go, this is a good section of code to pick up where you left off

#lastspp <- substr(fileList[length(fileList)],nchar(myPathToData)+2,nchar(fileList[length(fileList)])-4)
#for (i in (match(lastspp,allAnts$scientificName)+1):dim(allAnts)[1]){
#      tmp <- try(dBase <- aw_data(scientific_name = allAnts$scientificName[i])$data)
#      if(is.null(tmp)==FALSE){
#        dBase$scientific_name <- paste(dBase$genus,dBase$specificEpithet,sep=" ")
#        write.csv(dBase,paste(myPathToData,paste(allAnts$scientificName[i],"csv",sep="."),sep="/"),row.names=FALSE)
#      }
#      if(class(tmp)=="try-error"){
#        doOverTaxa <- c(doOverTaxa,i)
#      }
#    dBase <- data.frame()
#  }
```
The nice thing about write csv files after every quiery is that time-outs won't completely scrap your progress.

```{r All Ant Data}
mySuffix <- '.csv'
fileList <- list.files(myPathToData, full.names=TRUE) 
fileList <- fileList[grep(mySuffix,fileList)]
antData <- multipleCombine(fileList, ply = ldply)

# This is the list of species that exist that we didn't pull from AntWeb the first time around
leftOverTaxa <- allAnts$scientificName[(allAnts$scientificName%in%sort(unique(antData$scientific_name)))==FALSE]
```

There are some `r length(leftOverTaxa)` that Ant Base says exist, but we about which we don't have data from Ant Web. I've written the AntBase species list and the compiled data from AntWeb (some `r length(antData$scientific_name)` species) to csv files in the data folder. 