---
title: "AntWeb"
author: "KE LeVan"
date: "January 13, 2016"
output: html_document
---

# Exploring patterns of ant occurrence data in AntWeb
Now we have some data from AntWeb. I want to make a cool interface to visualize the data on this page using D3. To do that effectively, my .csv file is going to have to become a .geojson file.

### Setting up libraries, paths and useful functions
```{r Libraries,warning=FALSE}
myPath <- "~/GitHub/antweb"
myPathToData <- paste(myPath,"data",sep="/")

library(dplyr,warn.conflicts = FALSE,quietly = TRUE)
library(geojsonio,warn.conflicts = FALSE,quietly = TRUE)
```

### Getting the data
```{r Data}
antData <- read.csv(paste(myPathToData,"antData.csv",sep="/"),header = TRUE,stringsAsFactors = FALSE)
colnames(antData)[colnames(antData)=="minimumElevationInMeters"] <- "elevation"
antData <- antData[colnames(antData)!="url" & colnames(antData)!="catalogNumber" & colnames(antData)!="typeStatus" 
                   & colnames(antData)!="stateProvince" & colnames(antData)!="dateIdentified" 
                   & colnames(antData)!="specificEpithet" & colnames(antData)!="geojson.type"]
antData <- antData[is.na(antData$decimal_latitude)==FALSE & is.na(antData$decimal_longitude)==FALSE,]

```

### Converting the csv
```{r csv file conversion}
geo <- geojson_json(antData[1:2000,],lat = "decimal_latitude",lon = "decimal_longitude")
geojson_write(input = geo,file = paste(myPathToData,"small_ants.geojson",sep="/"))
```

### Subsetting the geojson
As it turns out, writing the ant points to a geojson is ENORMOUS, which I should have anticipated. If I'm going to use these geojson files to visualize the data, then I'm going to have to get more creative about it. Fundamentally, I want to examine the following:

1.  What areas are ants most reported as occurring on AntWeb? 
1.  Which parts of the globe are richest?
1.  How does that change when richness normalized by reporting intensity?
1.  Picking out just one species, where does that species occur?

Eventually, I'd like to do the same at the genus and subfamily taxon-levels. To do any of this, though, I'm going to have to subset the data or synthesize them.
```{r synthesized geojson}
# Make a grid of the world
lon <- seq(-179.5,179.5,1) # Possible longitude
lat <- seq(-89.5,89.5,1) # Possible latitude
size <- length(lon)*length(lat) # size of the df
grid <- data.frame(lat=rep(lat,length(lon)),
                   lon=sort(rep(lon,length(lat))),
                   numRecords=rep(NA,size),
                   sppRichness=rep(NA,size),
                   genRichness=rep(NA,size),
                   subfamRichness=rep(NA,size),
                   row.names = 1:size)

# Filling the grid with ant data
for (i in 1:size){
  antData %>% 
    filter(decimal_latitude>(grid$lat[i]-0.5),
           decimal_latitude<(grid$lat[i]+0.5),
           decimal_longitude>(grid$lon[i]-0.5),
           decimal_longitude<(grid$lon[i]+0.5)) -> ants
  if (dim(ants)[1]>0){
  grid$numRecords[i] <- dim(ants)[1]
  ants <- unique.data.frame(ants[colnames(ants)=="subfamily"|
                                   colnames(ants)=="genus"|
                                   colnames(ants)=="scientific_name"])
  grid$sppRichness[i] <- dim(ants)[1]
  ants <- unique.data.frame(ants[colnames(ants)=="subfamily"|
                                   colnames(ants)=="genus"])
  grid$genRichness[i] <- dim(ants)[1]
  ants <- unique.data.frame(ants[colnames(ants)=="subfamily"])
  grid$subfamRichness[i] <- dim(ants)[1]
  print((i/size)*100)
  }
}
rm(ants)

# Now that I have a grid, I can scale the occurence/diversity
# raw numbers to make an attractive scalar for my character 
grid$numRecords <- (grid$numRecords/max(grid$numRecords,na.rm = TRUE))*20
grid$sppRichness <- (grid$sppRichness/max(grid$sppRichness,na.rm = TRUE))*20
grid$genRichness <- (grid$genRichness/max(grid$genRichness,na.rm = TRUE))*20
grid$subfamRichness <- (grid$subfamRichness/max(grid$subfamRichness,na.rm = TRUE))*20
```

Now that I have summary stats for each grid location, I can create .geojson files for grids at each of my resolutions of interest.

```{r making smaller geojson files}
# Number of occurrences
geo <- geojson_json(grid[is.na(grid$numRecords)==FALSE,
                         colnames(grid)=="lat" |
                         colnames(grid)=="lon" |
                         colnames(grid)=="numRecords"],lat = "lat",lon = "lon")
geojson_write(input = geo,file = paste(myPathToData,"numRecords.geojson",sep="/"))

# Species Richness
geo <- geojson_json(grid[is.na(grid$sppRichness)==FALSE,
                         colnames(grid)=="lat" |
                         colnames(grid)=="lon" |
                         colnames(grid)=="sppRichness"],lat = "lat",lon = "lon")
geojson_write(input = geo,file = paste(myPathToData,"speciesRichness.geojson",sep="/"))

# Genus Richness
geo <- geojson_json(grid[is.na(grid$genRichness)==FALSE,
                         colnames(grid)=="lat" |
                         colnames(grid)=="lon" |
                         colnames(grid)=="genRichness"],lat = "lat",lon = "lon")
geojson_write(input = geo,file = paste(myPathToData,"genusRichness.geojson",sep="/"))

# Subfamily Richness
geo <- geojson_json(grid[is.na(grid$subfamRichness)==FALSE,
                         colnames(grid)=="lat" |
                         colnames(grid)=="lon" |
                         colnames(grid)=="subfamRichness"],lat = "lat",lon = "lon")
geojson_write(input = geo,file = paste(myPathToData,"subfamilyRichness.geojson",sep="/"))
```